---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kostra2010R

<!-- badges: start -->
<!-- badges: end -->

The main goal of kostra2010R is to provide access to KOSTRA-DWD-2010R dataset from within R.

Abstract from the [official description](https://opendata.dwd.de/climate_environment/CDC/grids_germany/return_periods/precipitation/KOSTRA/KOSTRA_DWD_2010R/gis/DESCRIPTION_gridsgermany_return_periods_precipitation_KOSTRA_KOSTRA_DWD_2010R_gis_en.pdf): 

_These vector data sets for GIS contain statistical precipitation values as a function of the duration and the return period. The scope of the data is the engineering dimensioning of water management structures. These include, sewerage networks, sewage treatment plants, pumping stations and retention basins. They are also often used for the dimensioning of drainage systems and infiltration systems. With the help of the data, however, it is also possible to estimate the precipitation level of severe heavy precipitation events with regard to their return periods. This estimation is often used to assess damage events._

_The data set contains the vector data sets of all 18 duration levels. A vecor data set contains the statistical precipitation (hN, design precipitation) of the present duration level D for nine return periods Tn (1-100 a) for the whole grid spanning 79 × 107 cells. INDEX_RC describes the unique identifier of a grid cell._

## Installation

You can install the development version of kostra2010R with:

```{r}
# install.packages("devtools")
# devtools::install_github("falk-env/kostra2010R")
```

Then load the package with

```{r}
library(kostra2010R)
```

## Usage

### Get "INDEX_RC" based on row and column information

Sometimes identification of grid cells is not accomplished using "INDEX_RC" directly but rather using a combination of X and Y information (e.g. column 11, row 49). This information can easily be used to generate the necessary "INDEX_RC" field.

```{r}
# Generate "INDEX_RC" out of X and Y information.
idx_build(11, 49)
```

If you wanted to check whether this constructed "INDEX_RC" field is really present in the data set (or you found an ID in some report and are not sure, if it is still being used), make use of the following function.

```{r}
# Is the following "INDEX_RC" entry present in the data set?
idx_exists("49011")
```

### Get "INDEX_RC" based on spatial information

The most common use case will be to get the relevant "INDEX_RC" based on coordinates provided, e.g. for the location of a precipitation station in order to be able to classify duration-specific precipitation heights in terms of return periods.

```{r}
# Sf objects created based on specified coordinates. Don't forget to pass the CRS.
p1 <- sf::st_sfc(
  sf::st_point(
    c(367773, 5703579)
  ),
  crs = 25832
)

p2 <- sf::st_sfc(
  sf::st_point(
    c(6.09, 50.46)
  ),
  crs = 4326
)

# Get indices by topological intersection between location point and grid cells.
idx_get(p1)
idx_get(p2)
```

### Cell-specific statistics from KOSTRA-DWD-2010R  

Now that we have messed a little with the grid cell identifiers, let's get a sneak peek into the data set itself based on the "INDEX_RC" specified.

```{r}
# Build a tibble containing precipitation heights as a function of duration and return periods for the grid cell specified.
kdata <- get_stats("49011")

kdata
```

The value of "INDEX_RC" has been assigned as an attribute to the tibble.

```{r}
# Attribute query
attr(kdata, "index_rc")
```

### Return period determination

Finally, we want to determine the return period according to the data set (without interpolating values at the moment) for a precipitation height and duration given.

```{r}
# Let's assume we measured 72.3 mm in 24 h ...
get_returnp(kdata, 72.3, 1440)
```

Accordingly, the approximate corresponding recurrence interval resp. annuality of this event amounts to something between 30 and 50 years as per KOSTRA-DWD-2010R.

### Further utilization

Data can now be visualized via `ggplot2` after a quick wide-to-long conversion...

```{r}
library(ggplot2)

# column name extraction for name/value junction
cnames <- colnames(kdata)[colnames(kdata) %>% stringr::str_detect("HN_*")]

# making use of tidyr
longdata <- tidyr::pivot_longer(kdata, cols = cnames)

# plot the whole data set, colors according to return periods
ggplot(longdata, aes(D_min, value, colour = name)) + 
  geom_point() +
  geom_line() +
  xlab("duration [min]") +
  ylab("precipitation height [mm]") +
  ggtitle("Precipitation height as a function of duration and return periods as per KOSTRA-2010R",
          subtitle = paste0("INDEX_RC: ", attr(kdata, "index_rc")))
```

... or exported to disk using `write.csv2()`.

## Contributing

See [here](https://github.com/rywhale/kiwisR/blob/master/.github/CONTRIBUTING.md) if you'd like to contribute.

## Legal information

According to the [terms of use](https://opendata.dwd.de/climate_environment/CDC/Terms_of_use.pdf) for data from GWS' CDC-OpenData area, reely accessible data may be re-used without any restrictions provided that the source reference is indicated, as laid down in the [GeoNutzV](http://www.gesetze-im-internet.de/geonutzv/index.html) ordinance.

Copyright: [Deutscher Wetterdienst](https://www.dwd.de/EN/service/copyright/templates_dwd_as_source.html)

## References

* Deutscher Wetterdienst > Leistungen > [KOSTRA-DWD](https://www.dwd.de/DE/leistungen/kostra_dwd_rasterwerte/kostra_dwd_rasterwerte.html)
* Deutscher Wetterdienst > OpenData > [KOSTRA-DWD raster data](https://opendata.dwd.de/climate_environment/CDC/grids_germany/return_periods/precipitation/KOSTRA/KOSTRA_DWD_2010R/)
