---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kostra2010R

<!-- badges: start -->
<!-- badges: end -->

The main goal of kostra2010R is to provide access to KOSTRA-DWD-2010R dataset 
from within R.

Official description: 

These vector data sets for GIS contain statistical precipitation values as a function of the duration and the return period. The scope of the data is the engineering dimensioning of water management structures. These include, sewerage networks, sewage treatment plants, pumping stations and retention basins. They are also often used for the dimensioning of drainage systems and infiltration systems. With the help of the data, however, it is also possible to estimate the precipitation level of severe haevy precipitation events with regard to their return periods. This estimation is often used to assess damage events.

The data set contains the vector data sets of all 18 duration levels. A vecor data set contains the statistical precipitation (hN, design precipitation) of the present duration level D for nine return periods Tn (1-100 a) for the whole grid spanning 79 × 107 cells. INDEX_RC describes the unique identifier of a grid cell.

## Installation

You can install the development version of kostra2010R from [GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
# devtools::install_github("falk-env/kostra2010R")
```

Then load the package with

```{r}
library(kostra2010R)
```

## Usage

### idx_exists()

Let's say you got a unique grid ID from some report and you want to check, if 
this is still present in the recent data set of KOSTRA-DWD.

```{r}
# Is the following "INDEX_RC" entry present in the data set?
idx_exists("49011")
```

Nice. You can be safe to generate statistics for this ID.

### build_idx()

Sometimes identification of the grid cells is not accomplished using "INDEX_RC"
but rather a combination of X and Y information (e.g. column 11, row 49). This
information can easily be used to generate the accordant "INDEX_RC" field.

```{r}
# Generate "INDEX_RC" out of X and Y information.
build_idx(11, 49)
```

### get_idx()

The most common use case will be to get "INDEX_RC" based on specific coordinates,
e.g. for the location of a precipitation station in order to be able classify 
measured precipitation heights.

```{r}
# Sf object created based on specified coordinates. Don't forget to pass the CRS.
p <- sf::st_sfc(
  sf::st_point(
    c(367773, 5703579)
    ), 
  crs = 25832
  )

p <- sf::st_sfc(
  sf::st_point(
    c(7.09, 51.46)
    ),
  crs = 4326
  )

# Topological intersection between location point and grid cells.
get_idx(p)
```

Ok, this seems to work also.

### build_table()

Now that we have messed a lot with the grid cell identifiers, let's get a sneak
peek into the data set itself based on "INDEX_RC".

```{r}
# Build a tibble containing precipitation heights as a function of duration and 
# return period for the grid cell specified.
build_table("49011")
```

The value of "INDEX_RC" has been assigned as an attribute to the tibble.

```{r}
# Create tibble
data <- build_table("49011")

attr(data, "index_rc")
```

### Further utilization

Data can now be visualized in form of plots...

```{r}
# Create tibble
data <- build_table("49011")

plot(data$D_min, 
     data$HN_100A_,
     xlab="interval [min]",
     ylab="precipitation height [mm]")
```

... or exported to disk using `write.csv2()`.

## Contributing

See [here](https://github.com/rywhale/kiwisR/blob/master/.github/CONTRIBUTING.md) if you'd like to contribute.

## Legal information

According to the [terms of use](https://opendata.dwd.de/climate_environment/CDC/Terms_of_use.pdf) for data from GWS' CDC-OpenData area, reely accessible data may be re-used without any restrictions provided that the source reference is indicated, as laid down in the [GeoNutzV](http://www.gesetze-im-internet.de/geonutzv/index.html) ordinance.

Copyright: [Deutscher Wetterdienst](https://www.dwd.de/EN/service/copyright/templates_dwd_as_source.html)

## References

* Deutscher Wetterdienst > Leistungen > [KOSTRA-DWD](https://www.dwd.de/DE/leistungen/kostra_dwd_rasterwerte/kostra_dwd_rasterwerte.html)
* Deutscher Wetterdienst > OpenData > [KOSTRA-DWD Rasterdaten](https://opendata.dwd.de/climate_environment/CDC/grids_germany/return_periods/precipitation/KOSTRA/KOSTRA_DWD_2010R/)
* Deutscher Wetterdienst > KOSTRA-DWD [Datensatzbeschreibung](https://opendata.dwd.de/climate_environment/CDC/grids_germany/return_periods/precipitation/KOSTRA/KOSTRA_DWD_2010R/gis/DESCRIPTION_gridsgermany_return_periods_precipitation_KOSTRA_KOSTRA_DWD_2010R_gis_en.pdf)
